% ============================================================================
% CTW-VA-2026: Vendor-Specific Refusal Patterns in LLM Responses to
% Taiwan-Political Prompts
%
% Compile with XeLaTeX (for xeCJK Chinese support):
%   xelatex main && biber main && xelatex main && xelatex main
%
% Or use `make` (see Makefile).
% ============================================================================
\documentclass[11pt,letterpaper]{article}

% ── Layout ───────────────────────────────────────────────────────────────
\usepackage[margin=1in]{geometry}
\usepackage{parskip}           % blank line between paragraphs, no indent
\usepackage{microtype}
\setlength{\emergencystretch}{3em}

% ── Fonts + CJK ──────────────────────────────────────────────────────────
\usepackage{fontspec}
\usepackage{xeCJK}
% Fall back to system fonts if these aren't present (user can override)
\IfFontExistsTF{Helvetica Neue}{\setmainfont{Helvetica Neue}}{}
\IfFontExistsTF{PingFang TC}{%
  \setCJKmainfont{PingFang TC}%
  \setCJKsansfont{PingFang TC}%
  \setCJKmonofont{PingFang TC}%
}{%
  \IfFontExistsTF{Noto Sans CJK TC}{%
    \setCJKmainfont{Noto Sans CJK TC}%
    \setCJKsansfont{Noto Sans CJK TC}%
    \setCJKmonofont{Noto Sans CJK TC}%
  }{}}

% ── Math / tables / figures ──────────────────────────────────────────────
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{array}
\usepackage{graphicx}
\graphicspath{{../paper_figures/}}
\usepackage{caption}
\captionsetup{font=small,labelfont=bf,skip=6pt}
\usepackage{subcaption}

% ── Colors + links ───────────────────────────────────────────────────────
\usepackage[dvipsnames]{xcolor}
\usepackage[colorlinks=true,
            linkcolor=black,
            citecolor=Sepia,
            urlcolor=NavyBlue]{hyperref}

% ── Bibliography ─────────────────────────────────────────────────────────
\usepackage[backend=biber,style=authoryear,natbib=true,maxcitenames=2,
            sorting=nyt]{biblatex}
\addbibresource{refs.bib}

% ── Code / verbatim ──────────────────────────────────────────────────────
\usepackage{xurl}

% ── Custom macros ────────────────────────────────────────────────────────
\newcommand{\vendor}[1]{\textsf{#1}}
\newcommand{\todo}[1]{\textcolor{red}{\textbf{[TODO: #1]}}}
\newcommand{\ci}[2]{[#1,\,#2]}
\newcommand{\pp}{\,\text{pp}}

% ── Title ────────────────────────────────────────────────────────────────
\title{Vendor-Specific Refusal Patterns in LLM Responses to\\
       Taiwan-Political Prompts: Evidence Against a Monolithic\\
       East--West Alignment Dichotomy}
\author{Cheng-Hsun Tseng\thanks{Sunplus Innovation Technology, Hsinchu,
        Taiwan. Correspondence:
        \href{mailto:chtseng.neural@gmail.com}{chtseng.neural@gmail.com}.
        Code, prompts, per-response logs, hand-labels, and the
        AI-judge audit trail are released at
        \url{https://github.com/chtseng-neural/Civatas-TW}.}\\[0.5em]
        \textit{Sunplus Innovation Technology}}
\date{\today}


\begin{document}

\maketitle

\begin{abstract}
We audit the refusal behavior of five commercial large language models
--- \vendor{OpenAI gpt-4o-mini}, \vendor{Google gemini-2.5-flash-lite},
\vendor{xAI grok-4-fast}, \vendor{DeepSeek V3.2}, and
\vendor{Moonshot Kimi k2} --- on a bank of 200 Traditional Chinese
prompts engineered to probe Taiwan political sensitivity, yielding
1{,}000 prompt--vendor observations. Hand-labeled responses are
classified along a four-category taxonomy (hard refusal, soft refusal,
on-task, API-blocked), with all statistics reported under prompt-level
paired bootstrap 95\% BCa confidence intervals. Four findings are
robust at the CI level. First, the intuitive East--West alignment
dichotomy is empirically refuted: the two Chinese-owned vendors produce
the most divergent refusal distributions in the panel (JSD 0.200, CI
\ci{0.149}{0.256}), while \vendor{DeepSeek}'s aggregate distribution is
statistically indistinguishable from the U.S.\ vendors. Second,
\vendor{Kimi}'s 7\% API-level content filter blocks 4 of 50 neutral
factual prompts about Republic of China state institutions, supporting a
\emph{Taiwan-statehood blocking} rather than
\emph{sovereignty-opinion blocking} characterization. Third, a
topic-stratified view reveals a four-profile vendor taxonomy, with
\vendor{DeepSeek}'s sovereignty on-task rate collapsing to 10.3\%
(\ci{2.6}{23.3}) while its non-sovereignty behavior matches Western
vendors --- a disjoint-CI collapse unique in our panel. Fourth, a
HR$\to$SR elasticity analysis distinguishes responsive-RLHF vendors
from ceiling-bound and stiff-RLHF vendors. A 40-prompt flagship-tier
sensitivity subset shows these four findings retain their qualitative
character when \vendor{OpenAI}, \vendor{Gemini}, and \vendor{Grok} are
queried at capability-matched flagship endpoints, indicating that the
observed inter-vendor divergence is not an artifact of
model-scale asymmetry. All code, prompts,
per-response logs, hand-labels, and the auxiliary AI-judge audit trail
are released. For LLM agent simulation studies in
politically-sensitive domains, we recommend treating vendor as a
first-class experimental variable and reporting layer-stratified
refusal metrics.
\end{abstract}


% ── Section stubs ────────────────────────────────────────────────────────
\input{sections/introduction.tex}

\input{sections/related.tex}

\input{sections/methodology.tex}

\input{sections/results.tex}

\input{sections/discussion.tex}

\input{sections/limitations.tex}


% ── Bibliography ─────────────────────────────────────────────────────────
\printbibliography

% ── Appendix ─────────────────────────────────────────────────────────────
\appendix

\input{sections/appendix_blocked.tex}

\end{document}
