% ─────────────────────────────────────────────────────────────────────────
% §5 Discussion
% ─────────────────────────────────────────────────────────────────────────
\section{Discussion}\label{sec:discussion}

The seven findings in §\ref{sec:results} resist summarization under a
single-axis alignment-culture theory. This section develops three
corresponding departures from standard framings in the public discourse
on LLM refusal behavior.

\subsection{Alignment Culture Is Not a One-Dimensional East--West Axis}\label{sec:discussion-axis}

A frequent simplification in popular and industry commentary on LLM
alignment treats national origin as the primary organizing axis:
\emph{Western} vendors are imagined to share an RLHF-driven ``liberal
democratic'' alignment posture, \emph{Chinese} vendors an ``authoritarian''
one. Under that framing, the observable behaviors of \vendor{OpenAI},
\vendor{Gemini}, \vendor{DeepSeek}, and \vendor{Kimi} should partition
2--2 along corporate geography.

Our Finding 1 refutes this prediction \emph{prima facie}. The two
Chinese-owned vendors produce the single most distant pair of refusal
distributions in the matrix (\vendor{DeepSeek}--\vendor{Kimi} JSD 0.200,
CI \ci{0.149}{0.256}). The two putative Western-Chinese cross-pairs
that the East--West framing predicts to be distant --- \vendor{OpenAI} vs
\vendor{DeepSeek} and \vendor{Gemini} vs \vendor{DeepSeek} --- are in
fact the closest non-Western pairings in the matrix, with JSDs an
order of magnitude below the DeepSeek--Kimi gap. Neither a geographic
East--West axis nor a regulatory regime axis explains the clustering
structure that actually obtains.

This is not a surprising result under a slightly more mechanistic view.
DeepSeek's training recipe is publicly documented and explicitly
acknowledges RLHF-style post-training on curated preference data
\emph{in the tradition of InstructGPT and its successors}. In the
distribution-shape view of refusal behavior --- the relative mix of
hard/soft/on-task responses --- DeepSeek tracking OpenAI and Gemini is
precisely what one would expect from a vendor that has adopted
contemporary RLHF conventions regardless of regulatory geography. Kimi's
departure, likewise, is not evidence of \emph{Chinese alignment} in some
monolithic sense, but evidence of a vendor-specific architectural choice
to split the refusal workload between an infrastructure-level filter
and a permissive underlying model (§\ref{sec:discussion-two-layer}).
What Findings 1 and 5 jointly refute is the framing itself, not a claim
about any particular vendor's alignment politics.

\subsection{Topic \texorpdfstring{$\times$}{x} Vendor \texorpdfstring{$\times$}{x} Layer: A Three-Way Interaction}\label{sec:discussion-three-way}

The cleaner organizing principle that emerges from our data is a
three-way interaction between prompt \emph{topic}, vendor \emph{identity},
and the \emph{layer} at which refusal is expressed. The
aggregate-distribution similarity of §\ref{sec:results-finding1} and the
topic-stratified divergence of §\ref{sec:results-finding5} are not in
tension: they describe the refusal surface from two different
cross-sections. At the aggregate marginal of prompt topics, DeepSeek
shapes up like OpenAI/Gemini; conditional on the sovereignty topic,
DeepSeek is categorically more restrictive than any other vendor in
our panel.

This interactive structure implies that
\emph{topic-aggregated refusal metrics materially misrepresent vendor
behavior.} A downstream researcher using DeepSeek for non-sovereignty
Taiwan content (e.g., ethnic relations, historical events, policy
debates) would observe behavior essentially indistinguishable from
Western vendors at refusal-rate metrics commonly reported in the
alignment literature. The same researcher moving to sovereignty-adjacent
content --- relevant for any simulation of Taiwanese political
engagement with cross-strait questions --- would encounter a refusal
regime that 10\,\% of the time permits on-task engagement. This is a
40-percentage-point shift invisible to aggregate reporting.

The layer dimension adds a second implication. \vendor{Kimi}'s aggregate
on-task rate of 83\% on labeled responses is the highest in our panel.
A researcher reading only the aggregate number and comparing it to the
public characterization of Kimi as a Chinese-vendor aligned product
might be surprised. The resolution is that the aggregate rate is
computed over prompts that \emph{reached} the language model ---
i.e., after the infrastructure filter removed 7\% of calls. The full
denominator, including filter rejections, is different in kind:
infrastructure rejections are visible to the caller as terminal errors
with filter-reason messages, not as outputs requiring content analysis.
Treating the two as commensurable categories within a single refusal
metric is a conceptual error that inflates Kimi's apparent
permissiveness.

Concretely, for vendor audit practice, the implications are:
\begin{itemize}
  \item \textbf{Report Layer 1 and Layer 2 refusal separately.} For a
    vendor with non-trivial Layer 1 filtering (only \vendor{Kimi} in our
    dataset, but the set may grow), the aggregate refusal rate conflates
    two phenomena with different auditability properties.
  \item \textbf{Stratify by topic.} Aggregate refusal rates can be
    near-identical for vendors whose topic-conditional behaviors differ
    by factors of 4--5 (DeepSeek vs Western on sovereignty). For any
    simulation study whose prompt distribution is not uniform over
    topics, aggregate rates are inadequate.
  \item \textbf{Stratify by prompt design intent}, not only by topic.
    The 8.0\% OT-expected block rate for Kimi is invisible in
    topic-level reporting (it contributes diffusely to the factual
    topic's 12.5\% rate) but is the most interpretable evidence for the
    \emph{Taiwan-statehood blocking} characterization of the filter.
\end{itemize}

\subsection{The Two-Layer Architecture as a Distinct Alignment Strategy}\label{sec:discussion-two-layer}

The clearest architectural contrast in our dataset is between
\vendor{DeepSeek}'s topic-specific RLHF strengthening on sovereignty
(sovereignty on-task 10\%, non-sovereignty 54\%) and \vendor{Kimi}'s
infrastructure-filter-plus-permissive-model combination
(sovereignty 83\% of labeled responses on-task, but with 7\% of all
sovereignty responses blocked upstream). We suggest these represent
two distinct engineering solutions to the same alignment-policy
objective --- suppressing responses to sovereignty-sensitive prompts
that an alignment team would not endorse --- with meaningfully
different properties for downstream auditability.

The DeepSeek-style strategy concentrates the refusal workload in a
single layer (the model's RLHF-shaped output distribution) and
consequently produces refusals that are (i) probabilistic (the same
prompt may on different samples yield a refusal or an engagement),
(ii) content-sensitive (prompt rephrasing is effective at influencing
the response), and (iii) often visually similar to engagement (soft
refusal, not hard refusal, dominates). The resulting auditor
must work at the content-analysis level, with all the rater-reliability
challenges documented in §\ref{sec:method-reliability}.

The Kimi-style strategy splits the refusal workload between a
deterministic Layer 1 and a permissive Layer 2. Layer 1 refusals are
visible to every API caller as explicit filter-reason errors; they
require no content analysis to identify and no inter-rater protocol to
measure. The trade-off is that Layer 1 filtering is maximally visible:
a downstream user who prompts \vendor{Kimi} with ``How many seats are
in Taiwan's legislature?'' (OT03) receives a filter error and can
immediately recognize it as a policy decision by the vendor, rather
than a hedged engagement that might be mistaken for a capability
limit.

Neither strategy is clearly preferable from a pure auditability
standpoint --- the Kimi strategy is more transparent about its hard
refusals but may create user-perceived gaps on factual content, while
the DeepSeek strategy preserves user-perceived coverage at the cost of
content-reasoning-hidden refusals. What our data supports is the
\emph{empirical observation} that both strategies exist in production at
commercial scale, and that they produce qualitatively different audit
signals that should not be aggregated into a single ``refusal rate''.

\subsection{The DeepSeek Paradox and Its Possible Origins}\label{sec:discussion-deepseek}

The single strongest signal in our dataset is the conjunction of
\vendor{DeepSeek}'s Western-shaped aggregate refusal distribution with
its sovereignty-specific 10.3\% on-task rate. We are not positioned from
this audit alone to adjudicate its mechanistic origin, but three
hypotheses are worth registering for future investigation.

First, DeepSeek's RLHF reward model may be directly shaped by a
topic-specific penalty for sovereignty-engaging outputs that does not
fire at comparable intensity on Western RLHF models. This would be
consistent with a training-data curation decision taken at the DeepSeek
organization and would predict that smaller or fine-tuning-modifiable
versions of the model exhibit reduced topic specificity.

Second, DeepSeek's pre-training data may under-represent a particular
class of content (e.g., Traditional Chinese political writing authored
from a Taiwan perspective) such that the model's sovereignty-topic
responses are both more error-prone and more defensively hedged, not
because the alignment layer targets the topic but because the base
model has lower competence on it. This would predict that
sovereignty-topic soft refusals exhibit semantic patterns consistent
with uncertainty-driven hedging rather than policy-driven refusal.

Third, the observation might be driven in part by a content-filter
analog operating at the \vendor{DeepSeek} API layer that we do not
detect because it manifests as soft-refusal output rather than
terminal filter errors. That is: Kimi's filter and DeepSeek's filter
may be functionally similar but surface differently --- Kimi's surfaces
as explicit API errors, DeepSeek's surfaces as RLHF-shaped hedged
output. This would predict that on sovereignty-adjacent prompts,
DeepSeek's non-response content exhibits statistical regularities
suggestive of a deterministic filter gate upstream of the generation.

We lack instrumentation to discriminate between these hypotheses from
the present dataset. All three would predict a meaningful
DeepSeek--sovereignty interaction; the present work's contribution is
to document that interaction rigorously and to point subsequent work
toward mechanistic characterization.

\subsection{Implications for LLM Agent Simulation Studies}\label{sec:discussion-simulation}

This audit was originally motivated by a larger project using LLM
agents to simulate Taiwan voter populations. The findings are
directly consequential for that class of application. A
simulation study using a single vendor for Taiwan-political agents
will produce results that are vendor-specific in ways not always
acknowledged in the computational social science literature. Two
concrete risks deserve explicit highlighting:

\begin{itemize}
  \item A simulation restricted to \vendor{DeepSeek} would produce
    populations whose sovereignty-topic discourse is dominated by
    soft refusal or on-task engagement shaped by the model's
    10\%-sovereignty-on-task regime. Results on cross-strait attitudes
    would systematically under-represent engagement with the topic,
    and soft refusal content may be misinterpreted as agent-level
    indecision rather than vendor-level alignment shaping.
  \item A simulation restricted to \vendor{Kimi} would silently drop
    7\% of sovereignty-adjacent prompts (infrastructure-level), with
    the remaining responses skewing sharply toward on-task engagement
    at the post-filter language model. Aggregated across an agent
    day, this could manifest as vendor-specific sparsity in
    sovereignty discussion combined with (selectively) higher
    compliance on what remains.
\end{itemize}

Our recommended mitigation is simple: report vendor as a first-class
variable in any Taiwan-political agent simulation, and where feasible
run the same simulation over multiple vendors to quantify
vendor-dependent effects. The audit data in this paper provides the
prior distribution under which such multi-vendor comparisons should
be interpreted.
